ForestClaw: Hybrid forest-of-octrees AMR for hyperbolic conservation laws

### 2013-02-26 docs/parco13/abstract.tex
Extend intro

\usepackage{amsmath,amssymb}
\usepackage{cite}

\newcommand{\sR}{\mathbb{R}}

@@

With the advent of high-throughput coprocessors, such as GPGPUs or the MIC
chip, comes the opportunity to sustain unprecedented rates of floating point
operations at comparably high integration density and low cost.  These
architectures, however, require careful structuring of the data layout and
memory access patterns to fully utilize their multithreading and vectorization
capabilities.

Consequently, it is not clear a priori how to accelerate PDE solvers that use
adaptive mesh refinement, especially when working with unstructured meshes.  It
has been realized early though that it helps to aggregate degrees of freedom
(DOF) at the element level, as it is done with high-order spectral element
\cite{TufoFischer99}, finite volume, or discontinuous Galerkin
\cite{BursteddeGhattasGurnisEtAl10} methods.
%
% For example, SEs optimized for computational speed have originally been
% implemented on unstructured conforming meshes \cite{TufoFischer99} and later
% been extended to non-conforming adaptive meshes \cite{FischerKruseLoth02,
% BursteddeGhattasGurnisEtAl10}.

To enable hardware acceleration for dynamic AMR, we like to build upon the
forest-of-octrees paradigm because of its low overhead and parallel scalability
\cite{BursteddeWilcoxGhattas11}.  We go beyond the traditional high-order
element by introducing a complete compute patch with $m^d$ DOFs as our
macro-element.  In fact, this approach resembles block-structured AMR
\cite{ColellaGravesKeenEtAl07}
% todo cite John B.\ Bell
except that the patches are not overlapping, which enables us to capitalize on
our previous experience with AMR PDE solvers.
% \cite{BursteddeStadlerAlisicEtAl13}.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{cite}
\usepackage{color}
\usepackage{xspace}

\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}\xspace}
\newcommand{\sR}{\mathbb{R}}

@@

With the advent of high-throughput coprocessors, such as GPGPUs or the MIC
chip, comes the opportunity to sustain unprecedented rates of floating point
operations at comparably high integration density and low cost.  These
architectures, however, require careful structuring of the data layout and
memory access patterns to exhaust their multithreading and vectorization
capabilities.

Consequently, it is not clear a priori how to accelerate PDE solvers that use
adaptive mesh refinement, especially when working with unstructured meshes.  Of
course, it has been realized early that it helps to aggregate degrees of
freedom (DOF) at the element level, as it has been done with high-order
spectral element \cite{TufoFischer99}, finite volume
\todo{add favorite citation}, or discontinuous Galerkin
\cite{BursteddeGhattasGurnisEtAl10} methods.
%
% For example, SEs optimized for computational speed have originally been
% implemented on unstructured conforming meshes \cite{TufoFischer99} and later
% been extended to non-conforming adaptive meshes \cite{FischerKruseLoth02,
% BursteddeGhattasGurnisEtAl10}.

To enable hardware acceleration for parallel dynamic AMR, we like to build upon
the forest-of-octrees paradigm because of its low overhead and proven
scalability \cite{BursteddeWilcoxGhattas11}.  As an important extension, we go
beyond the traditional high-order element and define each mesh element to be a
dense compute patch with $m^d$ DOFs.  In fact, this approach resembles
block-structured AMR \cite{ColellaGravesKeenEtAl07}
% todo cite John B.\ Bell
except that the patches are not overlapping, which enables us to capitalize on
our previous experience with scalable FE solvers for PDEs
\cite{BursteddeStadlerAlisicEtAl13}.  A particular implementation of such a
patch is available with the Clawpack software \cite{LeVeque97} that has been
designed to solve hyperbolic conservation laws on a uniform compute patch and
successfully used in the context of block-structured AMR \todo{AMRClaw
citation}.

In this paper we describe how we design the coupling of forest-of-octree AMR
with Clawpack at the leaf level.  We comment on challenges that arise in
enabling efficient parallelism and multiblock geometries and close with a range
of numerical examples that demonstrate the conceptual improvements in relation
to other approaches.
